Rate Transfers with Polyhedral Surrogate Losses /\\ Surrogate Regret Bounds for Polyhedral Losses

### 2021-05-19 rates-neurips-21/rates.tex
pause in Part 1 proof

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{fact}{Fact}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}

@@

\subsection{Property elicitation and the conditional approach}
Our technical approach is based on property elicitation, which studies the structure of $\ell$ and $L$ as they relate to $\simplex$.
This involves a \emph{conditional} perspective.
Given a data distribution $\D$ on $\X \times \Y$, for $(X,Y) \sim \D$, we consider the format of possible target predictions $g(x) = r \in \R$; surrogate predictions $h(x) = u \in \reals^d$; and conditional distributions $\Pr[ \cdot \mid X] = p \in \simplex$.
By dropping the role of the hypotheses and the feature space $\X$, we can focus on the relationships of the functions $\ell(r)_y$ and $L(u)_y$ on the conditional distribution $p$.

In particular, the \emph{Bayes risk} of $L$ is $\risk{L}(p) = \inf_{u \in \reals^d} \inprod{p}{L(u)}$, and similarly the Bayes risk of $\ell$ is $\risk{\ell}(p) = \inf_{r \in \R} \inprod{p}{\ell(r)}$.
We abuse notation to define the \emph{conditional surrogate regret} $R_L(u,p) = \inprod{p}{L(u)} - \risk{L}(p)$, and similarly the \emph{conditional target regret} $R_{\ell}(r,p) = \inprod{p}{\ell(r)} - \risk{\ell}(p)$.

To begin, we define a \emph{property}, which is essentially a statistic or summary of a distribution, such as the mode.
We then define what it means for a loss function to \emph{elicit} a property.
We write $\Gamma: \simplex \toto \R$ as shorthand for $\Gamma: \simplex \to 2^{\R} \setminus \{\emptyset\}$.
We write $\inprod{p}{L(u)}$ for the expected surrogate loss of prediction $u$ on distribution $p \in \simplex$, and similarly $\inprod{p}{\ell(r)}$ is the expected target loss of $r$.

@@

\bo{Define conditional transfer rate in this section?}

\section{Upper bound: polyhedral surrogates}
In this section, we develop the proof of Theorem \ref{thm:main-upper}.
At a high level, the proof has two parts:
\begin{enumerate}
  \item Fix $p \in \simplex$.
        Prove that $R_{\ell}(\psi(u),p) \leq \alpha_p R_L(u,p)$ for some constant $\alpha_p$.
  \item Apply this result to a finite subset $P$ of $\simplex$, namely the vertices of the level sets of $\gamma$.
        Argue (carefully) that, because a general $p$ is a convex combination of vertices of its own level set, $R_{\ell}(\psi(u),p) \leq \left(\max_{p' \in P} \alpha_{p'}\right) R_L(u,p)$.
        This implies a linear conditional transfer rate with constant $\max_{p' \in P} \alpha_{p'}$; it immediately implies a linear transfer rate as promised.
\end{enumerate}

\subsection{Part 1: fixed $p$}
\bo{include this proof sketch? Maybe not!}
\begin{enumerate}
  \item If $L$ is polyhedral, then the set of minimizers of $\inprod{p}{L(\cdot)}$ is a nice, convex, actually polyhedral subset of $\reals^d$.
  \item In addition, expected loss increases linearly with the distance of a prediction $u$ to this set.
  \item If $(L,\psi)$ is calibrated for $\ell$, then $\psi$ must be $\epsilon$-separated (Definition \bo{ref} below): all predictions with $\epsilon$ of the optimal set must link to an optimal target prediction in $\gamma(p)$.
  \item So conditional target regret is nonzero only for surrogate predictions that are $\epsilon$-far from the optimal set, which 
\end{enumerate}

>>>>

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{fact}{Fact}
\newtheorem{observation}{Observation}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}

@@

\subsection{Property elicitation and the conditional approach}
Our technical approach is based on property elicitation, which studies the structure of $\ell$ and $L$ as they relate to $\simplex$.
This involves a \emph{conditional} perspective.
Given a data distribution $\D$ on $\X \times \Y$, for $(X,Y) \sim \D$, we consider the format of possible target predictions $g(x) = r \in \R$; surrogate predictions $h(x) = u \in \reals^d$; and conditional distributions $\Pr[ \cdot \mid X] = p \in \simplex$.
By dropping the role of the hypotheses and the feature space $\X$, we can focus on the relationships of the functions $\ell(r)_y$ and $L(u)_y$ and the conditional distribution $p$.

In particular, the \emph{Bayes risk} of $L$ is $\risk{L}(p) = \inf_{u \in \reals^d} \inprod{p}{L(u)}$, and similarly the Bayes risk of $\ell$ is $\risk{\ell}(p) = \inf_{r \in \R} \inprod{p}{\ell(r)}$.
We abuse notation to define the \emph{conditional surrogate regret} $R_L(u,p) = \inprod{p}{L(u)} - \risk{L}(p)$, and similarly the \emph{conditional target regret} $R_{\ell}(r,p) = \inprod{p}{\ell(r)} - \risk{\ell}(p)$.
Given a rate transfer function $\zeta$, we say $(L,\psi)$ guarantees a \emph{conditional regret transfer rate} of $\zeta$ for $\ell$ if, for all $p \in \simplex$ and all $u \in \reals^d$, $R_{\ell}(\psi(u),p) \leq \zeta(R_L(u,p))$.

The following is a direct result of Jensen's inequality, using that $R_L(h;\D) = \E_X R_L(h(x),p_X)$ where $p_X$ is the conditional distribution on $Y$ given $X$.
\begin{observation} \label{obs:transfer}
  If $(L,\psi)$ guarantee a conditional regret transfer rate of $\zeta$ for $\ell$, and $\zeta$ is concave, then $(L,\psi)$ guarantee a regret transfer rate of $\zeta$ for $\ell$.
\end{observation}

\paragraph{Property elicitation.}
To begin, we define a \emph{property}, which is essentially a statistic or summary of a distribution, such as the mode.
We then define what it means for a loss function to \emph{elicit} a property.
We write $\Gamma: \simplex \toto \R$ as shorthand for $\Gamma: \simplex \to 2^{\R} \setminus \{\emptyset\}$.
We write $\inprod{p}{L(u)}$ for the expected surrogate loss of prediction $u$ on distribution $p \in \simplex$, and similarly $\inprod{p}{\ell(r)}$ is the expected target loss of $r$.

@@

\section{Upper bound: polyhedral surrogates}
In this section, we develop the proof of Theorem \ref{thm:main-upper}.
At a high level, the proof has two parts:
\begin{enumerate}
  \item Fix $p \in \simplex$.
        Prove that $R_{\ell}(\psi(u),p) \leq \alpha_p R_L(u,p)$ for some constant $\alpha_p$.
  \item Apply this result to a finite subset $P$ of $\simplex$, namely the vertices of the level sets of $\gamma$.
        Argue (carefully) that, because a general $p$ is a convex combination of vertices of its own level set, $R_{\ell}(\psi(u),p) \leq \left(\max_{p' \in P} \alpha_{p'}\right) R_L(u,p)$.
        This implies a linear conditional transfer rate with constant $\max_{p' \in P} \alpha_{p'}$; it immediately implies a linear transfer rate as promised.
\end{enumerate}
\bo{Not sure what to say here.}
We view both parts as relatively surprising.
The first contains several important technical steps, as we will outline next.
Intuitively, it is not clear why the first part should imply the main result: it might be that $\sup_p \alpha_p = \infty$, for example, by approaching a boundary.

\subsection{Part 1: fixed $p$}

We begin with the following key fact, a sort of ``pointwise'' conditional transfer bound.
\begin{proposition} \label{prop:fix-p}
  Suppose the surrogate and link $(L,\psi)$ are consistent for the target $\ell$.
  Then for any $p \in \simplex$, there exists $\alpha_p \geq 0$ such that, for all $u \in \reals^d$, $R_{\ell}(\psi(u),p) \leq \alpha_p R_L(u,p)$.
\end{proposition}

For intuition, observe that $R_{\ell}(\psi(u),p)$ takes on only finitely many values as we vary $p$, as the range $\R$ of $\psi$ is finite.
Therefore, it suffices to find a constant $C$ such that, if $R_{\ell}(\psi(u),p) > 0$, then $R_L(u,p) \geq C$.

\bo{include this proof sketch? Maybe not!}
\begin{enumerate}
  \item If $L$ is polyhedral, then the set of minimizers of $\inprod{p}{L(\cdot)}$ is a nice, convex, actually polyhedral subset of $\reals^d$.
  \item In addition, expected loss increases linearly with the distance of a prediction $u$ to this set.
  \item If $(L,\psi)$ is calibrated for $\ell$, then $\psi$ must be $\epsilon$-separated (Definition \bo{ref} below): all predictions with $\epsilon$ of the optimal set must link to an optimal target prediction in $\gamma(p)$.
  \item So conditional target regret is nonzero only for surrogate predictions that are $\epsilon$-far from the optimal set, which 
\end{enumerate}
